---
title: "Regression in Data Analysis Assignment"
date: "5/22/2019"
output: html_document
---

```{r setup, include=FALSE}
library('janitor')
library('lubridate')
library('tidyverse')
```

### Situation:  

A biotech company is researching a new plant-based ground beef alternative. They know how to increase the `sepal_length` of their plant but really need to know the effect on the `petal_length`.

Assume the data set is perfectly clean.

### Read in data
```{r}
dat = iris %>%
  transmute(sepal_length = Sepal.Length,
            petal_length = Petal.Length)
```

### Look at data
```{r}
head(dat)
```


### Plot the data  

Use `geom_point` and `geom_smooth` to visually identify any relationship between the two variables. Use `sepal_length` along the x-axis.
```{r}
dat %>%
  ggplot(...) +
  geom_point() + 
  geom_smooth()
```



### Pick slope and intercept values to create a model   

Keep changing the slope & intercept values to see what it does to your line of best fit. End with a line that seems to roughly identify the relationship between the `sepal_length` and `petal_length`
```{r}
# Equation for line of best fit
# y = m*x + b 
# m is slope, b is intercept
# residuals are the difference between y and the fit
# hint: petal_length is the `y` in this case, sepal_length is the `x`

m = 0  # Slope
b = 0  # Intercept

final_data = dat %>%
  mutate(fit = ...,
         residuals = ...,
         squared_residuals = residuals^2)


### The part below is done for you
final_data %>%
  ggplot(aes(x = sepal_length)) + 
  geom_point(aes(y = petal_length)) + 
  geom_line(aes(y = fit), col = 'red')
```


### Plot a histogram of the residuals  

In lay person terms, what do the residuals tell us and why should we care about them?
```{r}
final_data %>%
  ggplot(...) + 
  geom_histogram()
```


### Use the linear regression function  

Use the `lm` function to model your data, and then compare the results to the slope and intercept you chose above. In lay person terms, write a short description of what the slope of the model means.
```{r}
mod = lm(... ~ ..., data = dat)
summary(mod)
```

